Equalizer: A Mature Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Equalizers figure
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@17 db7c848b-7755-4df2-b763-8796797afa52

Equalizer are an addition to compound trees. They modify parameters of their
respective subtree at runtime to optimize one aspect of the decomposition. Due
to their nature, they are transparent to application developers, but might have
application-accessible parameters to tune their behaviour.
\subsubsection{Sort-First and Sort-Last Load Equalizer}

Sort-first and sort-last load balancing is the most obvious optimization for
these parallel rendering modes. Our load equalizer is fully transparent for
application developers, that is, it uses a reactive approach based on past
rendering times. This assumes a reasonable frame-to-frame coherency. Our
implementation stores a 2D or 1D grid of the load, mapping the load of each
channel. The load is stored in normalized 2D/1D coordinates using
$\frac{time}{area}$ as the load, the contributing source channels are organized
in a binary tree, and then the algorithm balances the two branches of each level
by equalizing the integral over the area on each side.

@@

Cross-segment load balancing addresses the optimal resource allocation of $n$
rendering resources to $m$ output channels (with $n\geq m$). The view equalizer
works in conjunction with load equalizer balancing the individual output
channels. It monitors the usage of shared source channels (across outputs) and
activates them to balance the rendering time of all outputs. In \cite{EEP:11},
we provide a detailed description and evaluation of our algorithm.

@@

The DFR equalizer provides a functionality similar to dynamic video resizing
\cite{MBDM:97}, that is, it maintains a constant framerate by adapting the
rendering resolution of a fill-limited application. In Equalizer, this works by
rendering into a source channel (typically on a FBO) separate to the destination
channel, and then scaling the rendering during the transfer (typically through
an on-GPU texture) to the destination channel. The DFR equalizer monitors the
rendering performance and accordingly adapts the resolution of the source
channel and zoom factor for the source to destination transfer. If the
performance and source channel resolutions allows, this will not only subsample,
but also supersample the destination channel to reduce aliasing artefacts.

@@

The monitor allows to reuse the rendering on another channel, typically for
monitoring a larger setup on a control workstation. Output frames on the display
channels are connected to input frames on a single monitoring channel. The
monitor equalizer changes the scaling factor and offset between the output and
input, so that the monitor channel has the same, but typically downscaled view,
as the originating segments.

>>>>

Equalizer are an addition to compound trees. They modify parameters of their
respective subtree at runtime to optimize one aspect of the decomposition. Due
to their nature, they are transparent to application developers, but might have
application-accessible parameters to tune their behaviour.

\begin{figure*}[ht]\center
  \subfigure[Load-Balancing]{\label{fig:loadeq}
    \includegraphics[width=0.24\textwidth]{images/loadeq}}\hfil
  \subfigure[Cross-Segment Load-Balancing]{\label{fig:vieweq}
    \includegraphics[width=0.24\textwidth]{images/vieweq}}\hfil
  \subfigure[Dynamic Frame Resolution]{\label{fig:dfr}
    \includegraphics[width=0.24\textwidth]{images/dfr}}\hfil
  \subfigure[Monitoring]{\label{fig:monitor}
    \includegraphics[width=0.24\textwidth]{images/monitoreq}}\hfil
  \caption{Runtime modifications}
  \label{fig:equalizers}
\end{figure*}

\subsubsection{Sort-First and Sort-Last Load Equalizer}

Sort-first (\fig{fig:loadeq}) and sort-last load balancing is the most obvious
optimization for these parallel rendering modes. Our load equalizer is fully
transparent for application developers, that is, it uses a reactive approach
based on past rendering times. This assumes a reasonable frame-to-frame
coherency. Our implementation stores a 2D or 1D grid of the load, mapping the
load of each channel. The load is stored in normalized 2D/1D coordinates using
$\frac{time}{area}$ as the load, the contributing source channels are organized
in a binary tree, and then the algorithm balances the two branches of each level
by equalizing the integral over the area on each side.

@@

Cross-segment load balancing (\fig{fig:vieweq}) addresses the optimal resource
allocation of $n$ rendering resources to $m$ output channels (with $n\geq
m$). The view equalizer works in conjunction with load equalizer balancing the
individual output channels. It monitors the usage of shared source channels
(across outputs) and activates them to balance the rendering time of all
outputs. In \cite{EEP:11}, we provide a detailed description and evaluation of
our algorithm.

@@

The DFR equalizer (\fig{fig:dfr}) provides a functionality similar to dynamic
video resizing \cite{MBDM:97}, that is, it maintains a constant framerate by
adapting the rendering resolution of a fill-limited application. In Equalizer,
this works by rendering into a source channel (typically on a FBO) separate to
the destination channel, and then scaling the rendering during the transfer
(typically through an on-GPU texture) to the destination channel. The DFR
equalizer monitors the rendering performance and accordingly adapts the
resolution of the source channel and zoom factor for the source to destination
transfer. If the performance and source channel resolutions allows, this will
not only subsample, but also supersample the destination channel to reduce
aliasing artefacts.

@@

The monitor equalizer (\fig{fig:monitor}) allows to reuse the rendering on
another channel, typically for monitoring a larger setup on a control
workstation. Output frames on the display channels are connected to input frames
on a single monitoring channel. The monitor equalizer changes the scaling factor
and offset between the output and input, so that the monitor channel has the
same, but typically downscaled view, as the originating segments.
